---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->

[![R-CMD-check](https://github.com/amymariemason/MR/actions/workflows/r.yml/badge.svg)](https://github.com/amymariemason/MR/actions/workflows/r.yml)
[![CRAN
status](https://www.r-pkg.org/badges/version/MendelianRandomization)](https://CRAN.R-project.org/package=MendelianRandomization)
[![R-CMD-check](https://github.com/amymariemason/MR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/amymariemason/MR/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


The MendelianRandomization R package provides a comprehensive set of methods for performing Mendelian randomization analyses with summarized data. We include methods for both single exposure (univariable) and multiple exposure (multivariable) analyses, as well as options to account for correlated variants in several methods. 

Summarized data on genetic associations with traits and diseases from large consortia can be accessed using websites such as [PhenoScanner](http://www.phenoscanner.medschl.cam.ac.uk/) and the [GWAS catalog](https://www.ebi.ac.uk/gwas/). These data can be used to obtain causal estimates following the Mendelian randomization paradigm. For an overview of this approach, see [here](https://www.mendelianrandomization.com/index.php/what-is-mr) for links to explanations and examples.

## Installation

```{r, eval = FALSE}
# Install released version from CRAN
install.packages("MendelianRandomization")
```

## Usage

If you are just getting started with Mendelian randomization, we recommend following the brief example analysis below, and then working through the package component vignettes, starting with the [data input vignette](https://amymariemason.github.io/MR/articles/Data_Input.html).

## Bugs and Feedback

If you encounter any issues, bugs, or have suggestions for improvements for our R package, please report to the package maintainer [Stephen Burgess](mailto:sb452@medschl.cam.ac.uk).

We are keen to include further summarized data Mendelian randomization methods in our package, and any other customized code that you want to share with others. You can [fork the package from Github](https://github.com/cran/MendelianRandomization) to test for compatibility with the existing package.

## Example analysis

This is an example analysis to investigate lipid effects on coronary artery disease risk using the example data included with the package.

```{r Image1}
library(MendelianRandomization) # load package

MRInputObject <- mr_input(bx = ldlc, 
                          bxse = ldlcse, 
                          by = chdlodds, 
                          byse = chdloddsse)
```

The variables *ldlc*, *hdlc*, *trig*, and *chdlodds* are the genetic associations with (respectively) LDL-cholesterol, HDL-cholesterol, triglycerides, and coronary heart disease (CHD) risk for 28 genetic variants reported by Waterworth et al (2010). The respective standard errors of the associations are given as *ldlcse*, *hdlcse*, *trigse*, and *chdloddsse*.

### Perform inverse-variance weighted method

We can use this package to perform a two-sample Mendelian Randomization analysis using a wide variety of methods. Here we show the inverse-weighted variance method:

```{r Image2, echo=TRUE}
MRAllObject_ivw <- mr_ivw(MRInputObject) # perform IVW method

MRAllObject_ivw  # view results

mr_plot(MRInputObject, line="ivw", interactive = FALSE) # plot results

```

### Alternative analysis methods

To assess the robustness of our results, we can perform other methods, such as the MR-Egger and weighted median methods:

```{r Image3}
MRAllObject <- mr_allmethods(MRInputObject, method="main") # perform MR-Egger and median methods

MRAllObject  # view results

mr_plot(MRAllObject) # plot results

```

For more methods, see the vignettes on [univariable](https://amymariemason.github.io/MR/articles/Univariable_MR_Methods.html) and [multivariable](https://amymariemason.github.io/MR/articles/Multivariable_MR_Methods.html) methods.
